---
title: "Data607 - Week3"
author: "Baron Curtin"
date: "February 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Environment Setup

```{r env, echo=FALSE}
library(tidyverse)
library(magrittr)
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
names <- raw.data %>%
  str_extract_all('[[:alpha:]., \\.]{2,}') %>%
  unlist %>%
  print
phone <- raw.data %>%
  str_extract_all("\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}") %>%
  unlist %>%
  print
```

## 1.) Rearrage the vector so all the elements conform to first_name last_name.
```{r one}
# remove salutations, prefixes, swap positions, and trim
swappedNames <- names %>%
  str_replace("[A-z]+\\.\\s", " ") %>%
  str_replace("(\\w+),\\s+(\\w+)", "\\2 \\1") %>%
  str_trim %>%
  print

# put data in data frame for easier viewing
df <- data_frame(swappedNames = swappedNames, phone = phone) %>%
  print
```

## 2.) Logical vector indicating whether a character has a title
```{r two}
# detect titles
title <- names %>%
  str_detect("[A-z]{2,3}\\.\\s*")

# put names and true/false vector in data frame
titles <- data_frame(names = names, titles = title) %>%
  print
```

## 3.) Logical vector indicating whether a character has a second name
```{r three}
# detect second name
twoNames <- names %>%
  str_detect("[A-Z]{1}\\.")

# put names and true/false vector in data frame
twoNameDf <- data_frame(names = names, twoNames = twoNames) %>%
  print
```

## 4.) Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.
* a.) [0-9]+\\$ - Select digits 0-9 one or more times, followed by a $
* b.) \\b[a-z]{1,4}\\b - Word boundaries, [a-z] 1-4x, Word boundaries
* c.) .*?\\.txt$ - Select files/file paths that have a text extension
* d.) \\d{2}/\\d{d}/\\d{4} - 2 digits/2 digits/4 digits (date format)
* e.) <(.+?)>.+?</\\1> - Used to find HTML code
```{r four}
qs <- c("4a", "4b", "4c", "4d", "4e")
test_str <- c("4$",
              "baron curtin bac curtin,baron",
              "C:/Users/ProblemX/OneDrive/Documents/Cuny SPS/Data 607/Week3/helloText.txt",
              "02/13/2018",
              "<span>Span tag</span>")
pats <- c("[0-9]+\\$", 
          "\\b[a-z]{1,4}\\b",
          ".*?\\.txt$",
          "\\d{2}/\\d{d}/\\d{4}",
          "<(.+?)>.+?</\\1>")

data_frame(patterns = pats, testStrings = test_str) %>%
  print
```

## 9.) The following code hides a secret message
```{r nine}
code <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

msg <- code %>%
  str_extract_all("[[A-Z].]") %>%
  unlist %>%
  str_replace_all("[.]", " ") %>%
  paste(collapse = "") %>%
  print
```

